---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data
The BRFSS data is collected through telephone interviews that take place between two sample: land line and cell phone calls. Phone numbers are collected from the CDC and randomly selected. Becuase of the data collects method (phone calls) and the randomization in sample selection, the data and subsequent conclusion in this analysis can be said to be generalized to the general population of phone owners. However, seeing as the data collected in the BRFSS is observational, only correlations can be infered. Nothing causitory can be draw. 


* * *

## Part 2: Research questions

**Research question 1:**
How does the mental health of those who are married compare to those who are unmarried? Marriage is often idealized as a chance at "happily ever after" and if this mainstream assumption holds true, we would expect a lower number of depression diagnosis among married populations.

independent variable: marital status (MARITAL)
dependent variables: depression diagnosis (ADDEPEV2) and number of days per year with poor mental health (MENTHLTH)

**Research question 2:**
Which states have the highest instances of skin cancer? Skin cancer is more likely to occur when we are exposed to UV radiation from the sun and improperly protected from it. With this in mind we may expect states with the highes UV indexes (Hawaii, Arizona, California, Florida, and Nevada) to have the highest instances of skin cancer. However, if this does not occur, we can infer that proper protection (sunscreen habits) may play a more important role than exposure. 

independent variable: state of residence (_STATE)
dependent variable: skin cancer rates (CHCSCNCR)

**Research question 3:**
How do exercise habits change as hours per week working change? Frequent exercise is essential for maintaining health across a lifetime, however, many American's find it difficult to keep a continuous exerciser routine. Often, this is attributed to laziness ,however, for many American's, work takes up a large part of their lives and energy expenditure. This part of the analysis aims to guage how hours worked impact hours of excersise 

independent variable: hours per week worked (SCNTWRK1)
dependent variable: Excersie outside of work (_TOTINDA)
* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research question 1:**

```{r Question 1: Numerical Summary}
married_count_dep_df <- brfss2013 %>%
  filter(!is.na(marital), !is.na(addepev2)) %>%
  group_by(marital_status = marital, depression_diagnosis = addepev2) %>%
  summarise(count = n())

marries_perc_dep_df <- brfss2013 %>% 
  filter(!is.na(marital), !is.na(addepev2)) %>%
  group_by(marital) %>%
  summarise(percent_with_depression_diagnosis = (sum(addepev2 == "Yes") / n()))

sixty_five_perc_by_age_df <- brfss2013 %>% 
  filter(!is.na(marital), !is.na(X_ageg5yr)) %>%
  group_by(marital == "Widowed") %>% 
  summarise(sixty_five_or_older = (sum(X_ageg5yr == "Age 65 to 69", X_ageg5yr == "Age 70 to 74 
", X_ageg5yr == "Age 75 to 79", X_ageg5yr == "Age 80 or older" )/ n()))
```
The above code first creates two data frames: The first one shows the counts for individuals with a depression diagnosis per martial status and the second one shows the percent of individuals with a depression diagnosis by marital status. From here we are able to see that the groups with the highest rates of depression diagnoses are individuals who are either separated or divorced. We can see those with the lowest rate of depression diagnoses are those who are either married or widowed. I found it a little shocking that those who are widowed have the second lowest rate of depression diagnosis. What could be happening here? If married individuals have lower rates of depression, couldn't we expect those with a dead spouse to be on the opposite end of the spectrum? 

This made me think of Laura Carstensen's research on older people, finding that these populations are among one of the happiest. Could this be a factor? The third data frame finds that, of those who are widowed, 66.9% of them are 65+. The lower rates of depression are likely a reflection of increased positive affect that occurs in older adults. 


```{r Question 1: Visual Summary}
levels(brfss2013$marital) <- gsub(" ", "\n", levels(brfss2013$marital))
ggplot(subset(brfss2013, !is.na(marital) & !is.na(addepev2)), aes(x = marital, fill = addepev2)) + geom_bar()

levels(marries_perc_dep_df$marital) <- gsub(" ", "\n", levels(marries_perc_dep_df$marital))
ggplot(marries_perc_dep_df, aes(x = marital, fill = percent_with_depression_diagnosis)) +geom_bar()

```
Graph #1 above shows the number of individuals with and without a depression diagnosis for each marital status. This graph helps intuit the rate of depression for each group. However, because the group sizes vary greatly, it can still be difficult to understand the exact rates. Graph #2 better visualizes the rate of depression per each marital group, with lighter colors indicatin high rates. Here we can see separated and divorced population have the highest rates of depression and married and widowed population have the lowest.

**Research question 2:**

```{r Question 2: Numerical Summary}
skin_cancer_rates_df <- brfss2013 %>%
  filter(!is.na(X_state), !is.na(chcscncr)) %>%
  group_by(X_state) %>%
  summarise(rate_of_skn_cnr = (sum(chcscncr == "Yes")/ n())) %>% 
  arrange(desc(rate_of_skn_cnr))

cant_see_dr_per_df <- brfss2013 %>%
  filter(!is.na(X_state), !is.na(checkup1)) %>%
  group_by(X_state) %>%
  summarise(cant_see_dr = (sum(checkup1 == "Within past 5 years", checkup1 == "5 or more years ago")/ n())) %>%
  arrange(desc(cant_see_dr))
```

This first data frame above find the states that have the highest instances of skin cancer. As expected, Florida and Arizona have the highest rates. California, Puerto Rico, and Guam are lower on the rankings which is surprising considering all three places have high UV indexes. One possible explanation is access to health care in these regions may be low. To investigate this, the second data frame looks at the rate of individuals within each region who have been unable to see a doctor within the last 2 year, due to cost. This data frame shows that lack of access may be a factor for people from California (9.4% unable to see a doctor) and Guam (7.8%), implying that there may individuals in those areas with skin cancer but no ability to recieve a diagnosis. 


```{r Question 2: Visual Summary}
ggplot(skin_cancer_rates_df, aes(x = X_state, y = rate_of_skn_cnr)) + geom_col() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

**Research question 3:**

```{r Question 3: Numerical Summary}


hours_worked_v_excersie_df <- brfss2013 %>% 
  filter(!is.na(scntwrk1), !is.na(X_totinda), !(scntwrk1 == 97)) %>%
  mutate(scntwrk1 = ifelse(scntwrk1 == 98, 0, scntwrk1), hours_worked_range = cut(scntwrk1, c(-Inf, 0, 10, 20, 30, 40, Inf), labels = c("0","1-10", "11-20", "21-30", "31-40", "40+"))) %>%
  group_by(hours_worked_range) %>%
  summarise(percent_work_out = (sum(X_totinda == "Had physical activity or exercise")  / n()))
```

The above code block creates a data frame that organizes cases into groups based on hours worked per week and calculates the percentage of members from each group that worked out within the last 30 days. Each case falls into one of the following groups: "0","1-10", "11-20", "21-30", "31-40", "40+". This data frame was achieved by removing 'NA's from the column capturing the hours worked per week and the column capturing if the case excersied within the last 30 days. Next, a column is added to the data frame to indicate which of the groups each case falls in. Finally, the data is grouped accordingly and the percentage of those who worked out in each group is calculated. 


```{r Question 3: Visual Summary}

ggplot(hours_worked_v_excersie_df, aes(x = hours_worked_range, y = percent_work_out)) + geom_col()


```


The above bar graph shows what percentage of individuals within each group had exercised with in the last 30 days. The group with the most individuals working out at least one within the last 30 days was the group that worked 40+ hours a week. The group who had the least number of individuals working out was 0 hours worked per week group. This runs counter to my intuition and I would like to do further analysis with data that shows the average amount of hours spent exercising by each group as opposed to tallying how many people worked out at least once with in the last 30 days. 

